---
title: "Diversity analysis in R"
author: "Ana Romani"
date: "2023-09-07"
output: html_document
---

```{r}
library(qiime2R)
library(phyloseq)
library(ggplot2)
library(dplyr)
```


```{r}
# Go to your directory that contain the qza files
dir <- "C:/Users/Romani Vasquez/Documents/Teach_notes/BioInfo/practice"
setwd(dir)
```

# Exploring data
```{r}
SVs <- read_qza("bespoke-taxonomy.qza") 
names(SVs)
table_features <- SVs$data
```

```{r}
taxon_frequency = as.data.frame(table(table_features$Taxon))
colnames(taxon_frequency)= c("Taxon","Frequency")
ggplot(data= taxon_frequency,aes(x=Taxon,y=Frequency))+ geom_histogram(stat = "identity")
ggplot(data= taxon_frequency,aes(x=Frequency))+ 
  geom_histogram(bins=50)+
  xlim(0,100)
```
# Generating a phylo seq object
```{r}
physeq_O <- qza_to_phyloseq(
  features = "table-deblur.qza",
  taxonomy = "bespoke-taxonomy.qza",
  metadata = "metadata.tsv"
)
```
# Estimate alpha diversity index based on phylo seq object
```{r}
richness <- estimate_richness(physeq_O)
richness$group = as.factor(ifelse(grepl(".C",rownames(richness)),"Children","Mother"))
```
You can also embed plots, for example:

```{r}
ggplot(data = richness,aes(x=group,y=Observed,color=group))+geom_boxplot()
```

```{r}
chao <-ggplot(data = richness,aes(x=group,y=Chao1,color=group))+geom_boxplot()+xlab("")
shannon <- ggplot(data = richness,aes(x=group,y=Shannon,color=group))+geom_boxplot()+xlab("")
simpson <- ggplot(data = richness,aes(x=group,y=Simpson,color=group))+geom_boxplot()+xlab("")
ggpubr::ggarrange(chao,shannon,simpson,ncol = 3,nrow = 1,common.legend = T)
```
# Significant test
The indexes have a significant difference?
## Check the distribution
```{r}
histogram(richness$Shannon)
#qqplot(richness$Shannon)
```
# No parametric test
# Mann U withney test
```{r}
children <- richness %>% filter(group=="Children")
children <- as.vector(children$Shannon)
mother <- richness %>% filter(group=="Mother")
mother <- as.vector(mother$Shannon)
wilcox.test(children,mother)
```

# Beta diversity

```{r}
library(vegan)
## OTU-based metrics
OTUs = physeq_O@otu_table@.Data #To use the “Group” column as the row names so that it will match our metadata
OTU_final = as.data.frame(t(OTUs)) 
BC.dist= as.matrix(vegdist(OTU_final, distance="bray"))
```

```{r}
### Ordination scatterplots (Bray-Curtis metric)
BC.nmds = metaMDS(OTU_final, distance="bray", k=2, trymax=20) 
# check the stress is lower than 0.2
```

```{r}
metadata <- read.csv("metadata.tsv",sep = "\t")
colnames(metadata)[1]<-"SampleID"
metadata$Group = as.factor(ifelse(grepl(".C",metadata$SampleID),"Children","Mother"))
data_plot <- as.data.frame(BC.nmds$points)
data_plot$SampleID = rownames(data_plot)
data_plot <- merge(metadata[,c(1,5)],data_plot,by="SampleID")
ggplot(data=data_plot,aes(x=MDS1,y=MDS2,color=Group)) + geom_point()
```
```{r}
#### Plot the nMDS
par(mfrow = c(1, 1))
##### Create a blank plot for the NMDS
plot(BC.nmds, type="n", main="Bray-Curtis")
##### Add the points colored by age
points(BC.nmds, pch=20, col=c("skyblue", "pink")[metadata$Group])
##### Add a legend
legend(0.2,-0.04,legend=c("Children","Mother"), col=c("skyblue","pink"), pch=20)
```

# Permutational multivariate analysis
```{r}
adonis2(BC.dist ~ Group, data = metadata, permutations = 100)
```

# Taxonomic ranges
```{r}
taxonomy <- read_qza("bespoke-taxonomy.qza")$data %>% parse_taxonomy()
freq_genus <- as.data.frame(table(taxonomy$Genus))
taxasums <- summarize_taxa(OTUs, taxonomy)$Genus
metadata <- read.csv("metadata.tsv",sep = "\t")
colnames(metadata)[1]<-"SampleID"
metadata$Group = as.factor(ifelse(grepl(".C",rownames(richness)),"Children","Mother"))
```

```{r}
library(reshape2)
```
```{r}
taxasums_t <- as.data.frame(t(as.matrix(taxasums)))
taxasums_t$SampleID <- row.names(taxasums_t)
all_tab <- merge(metadata[,c(1,5)],taxasums_t,by="SampleID")
long_all_tab <- melt(all_tab,id=c("SampleID","Group"))
```

```{r}
long_child_tab_10 <-long_all_tab %>% 
  filter(Group=="Children")%>%
  arrange(desc(value))%>%
  slice_head(n=10)
long_mother_tab_10 <-long_all_tab %>% 
  filter(Group=="Mother")%>%
  arrange(desc(value))%>%
  slice_head(n=10)
long_all_tab_10 <- rbind(long_child_tab_10,long_mother_tab_10)
ggplot(data=long_all_tab_10,aes(x=Group,y=value,fill=variable))+
  geom_bar(stat = "identity")+
  labs(fill = "Taxonomic range")+
  theme(legend.position = "right",
        legend.text = element_text(size = 5),
        legend.box = "vertical")
```
# Generating a phylo seq object
```{r}
physeq_O <- qza_to_phyloseq(
  features = "table-deblur.qza",
  tree = "insertion-tree.qza",
  taxonomy = "bespoke-taxonomy.qza",
  metadata = "ecam-sample-metadata.tsv"
)
```

```{r}
library(ggpubr)
a_my_comparisons <- list( "n", "y")

plot_richness(physeq_O, x = "antiexposedall", color = 'antiexposedall', measures = c("Observed", "Chao1","Shannon"), title = "Alpha diversity indices of samples") + 
  geom_boxplot(alpha = 0.6) +
  theme(legend.position = "right", axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=12))
```
```{r}
a_my_comparisons <- list( "Vaginal", "Cesarean")

plot_richness(physeq_O, x = "delivery", color = 'delivery', measures = c("Observed", "Chao1","Shannon","Simpson"), title = "Alpha diversity indices of samples") + 
  geom_boxplot(alpha = 0.6) +
  theme(legend.position = "right", axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=12))
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
